### 2006-03-08 papers/slides-TYPES'06/abstract.tex
Forgot to actually add the files

\def\myabstract{We set out to establish a phase distinction between writing and
  proving algorithms in the \Coq{} environment. Concretely, this means
  allowing to write algorithms as easily as in a pratical functionnal programming language
  whilst giving them as rich a specification as desired and proving that
  the code meets the specification using the whole \Coq{} proof
  apparatus. This is achieved by extending conversion to an equivalence which relates types
  and subsets based on them, a technique  
  originating from the ``\emph{Predicate subtyping}'' feature of \PVS{} and
  following mathematical convention. The typing judgments can be
  translated to \CCI{} by means of an interpretation which inserts
  coercions at the appropriate places. These coercions can contain
  existential variables representing the propositional parts of the final
  term. We will present a prototype implementation of this process
  which can be used in a variety of situations in the proof assistant.}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: