### 2006-05-31 papers/slides-ICFP'06/abstract-only.tex
Abstract for ws on metatheory

\documentclass{article}
\usepackage[latin1]{inputenc}
\usepackage{xspace} % To get the right spacings in front of : and so on
\usepackage[english]{babel}
\usepackage{abbrevs}
\usepackage{subfigure}
\usepackage{ifthen}
\usepackage{coqdoc}
\usepackage{bussproofs}
\usepackage{pgf,pgfarrows,pgfnodes}

\title{\Russell's metatheoretic study}
\date{}
\begin{document}
\maketitle
\abstract
\input{abstract}
\myabstract
\end{document}

%%% Local Variables: 
%%% mode: latex
%%% LaTeX-command: "x=pdf; TEXINPUTS=\"..:../style:../figures:\" ${pdfx}latex"
%%% TeX-master: t
%%% End: