### 2006-05-31 papers/slides-ICFP'06/abstract.tex

\def\myabstract{
  We are working on the formalization of \Russell{}'s type theory in the
  \Coq{} proof assistant. \Russell{} is a type system based on the
  Calculus of Constructions which is extended by an equivalence on types
  which include $\beta$-conversion. The extension permits to identify
  types and subsets based on them in a manner similar to the
  \emph{Predicate Subtyping} feature of \PVS{}. We are aiming at 
  a complete proof of \Russell{} metatheoretic properties (Subject
  Reduction, maybe Strong Normalization), the refining steps which
  led us to the typing algorithm and also the correctness of a
  traduction from \Russell{} to the Calculus of Inductive Constructions.
  The proof poses several technical difficulties including:
  \begin{itemize}
  \item Elimination of transitivity in a system with an untyped type
    conversion relation ;

  \item Subject Reduction, which is not directly provable for the
    declarative system. Here we adapted the technique developped by
    Robin Adams \cite{adamstposr}. It includes a new term algebra, with
    associated reduction operations and a new type system (actually two)
    for which we have to prove metatheoretic properties.

  \item Correction of the traduction: the target system include
    metavariables, introducing a second, unusual kind of variable binding.
  \end{itemize}

  We will present these difficulties and our recipes for solving them in
  the \Coq environment.
}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: