### 2007-03-02 papers/seminar-proval/abstract.tex
Prepare new slides

\def\myabstract{
  There exist different methods for developing certified programs in the \Coq
  environment. The standard one is to use \Coq's language as a purely
  functional version of \ML using simple types and prove properties
  about them in the usual way. Another, lesser known approach, is to use the
  complete \Coq language to specify and implement programs, making
  invariants, pre- and post-conditions appear in the types of terms. 
  This method is generally overlooked because it is
  difficult to cope with in practice as algorithms and proofs become
  entangled. \Program tries to solve this problem by reconciling the two
  approaches. It consists of a
  language (\Russell) which permits writing purely algorithmical code,
  a sound translation from \Russell terms to \Coq terms which
  generates proof obligations and a mechanism for handling these.
  I will present these three components then delve into an example of
  using \Program to build a certified implementation of the Finger Tree
  data structure of Hinze \& Patterson.

}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: