### 2007-04-01 papers/slides-TYPES'07/abstract.tex
TYPES'07 slides directory

\def\myabstract{
  We report on an experiment of using \Russell to develop complex
  programs in the \Coq proof assistant. We present a certified
  implementation of the
  Finger Tree data structure (Hinze \& Patterson, 2006) using the advanced techniques of
  specification available through the \Russell language, notably non-uniform
  inductive parameters and polymorphic recursion, subset specifications
  for datatypes and functions and dependent inductive types. Along the
  way, we show the respective merits and deficiencies of these
  techniques in our particular setting and discuss possible solutions.
  Finally we present an original result pertaining to the dependent
  specification style, giving us theorems for free.
  This development shows that programs with rich specifications
  can be written much like their original version (i.e. \Haskell)
  but the matter is how to \emph{reason} about them easily. 
  There is still room for improvement on this part.
}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: